---
title: "Ramji_Lab2"
author: "Anna Ramji"
date: 
output: pdf_document
---

```{r}
library(tidyverse)
library(stats)
library(tidymodels)
library("janitor")
library("corrplot")
dat <- read_csv(file = "https://raw.githubusercontent.com/MaRo406/eds-232-machine-learning/main/data/pumpkin-data.csv")
```


Today we will be continuing the pumpkin case study from last week. We will be using the data that you cleaned and split last time (pumpkins_train) and will be comparing our results today to those you have already obtained. Open and run your Lab 1.Rmd as a first step so those objects are available in your Environment.

Once you have done that, we'll start today's lab by specifying a recipe for a polynomial model.  First we specify a recipe that identifies our variables and data, converts the package variable to a numerical form, and then adds a polynomial effect with `step_poly()`

```{r}
# Specify a recipe
poly_pumpkins_recipe <-
  recipe(price ~ package, data = pumpkins_train) %>%
  step_integer(all_predictors(), zero_based = TRUE) %>% 
  step_poly(all_predictors(), degree = 3)
```

How did that work? Later we will learn about model tuning that will let us do things like find the optimal value for degree.  For now, we'd like to have a flexible model, so we'll use a relatively large value.

Polynomial regression is still linear regression, so our model specification looks similar to before.

```{r}
# Create a model specification called poly_spec
poly_spec <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```
Question 1: Now take the recipe and model specification that just created and bundle them into a workflow called `poly_df`.

```{r}
# Bundle recipe and model spec into a workflow
poly_wf <- workflow() |> 
  add_recipe(poly_pumpkins_recipe) |> 
  add_model(poly_spec)
```

Question 2: fit a model to the pumpkins_train data using your workflow and assign it to poly_wf_fit
```{r}
# Create a model
poly_wf_fit <- fit(data = pumpkins_train, object = poly_wf)
```

```{r}
# Print learned model coefficients
poly_wf_fit
```


```{r}
# Make price predictions on test data
poly_results <- poly_wf_fit %>% predict(new_data = pumpkins_test) %>% 
  bind_cols(pumpkins_test %>% select(c(package, price))) %>% # bind multiple dfs by these columns
  relocate(.pred, .after = last_col()) # put prediction column after price column for easier comparison

# Print the results
poly_results %>% 
  slice_head(n = 10)
```

Now let's evaluate how the model performed on the test_set using yardstick::metrics().
```{r}
metrics(data = poly_results,
        truth = price,
        estimate = .pred)
```

```{r}
metrics(data = lm_results,
        truth = price,
        estimate = .pred)
```

Question 3: How do the performance metrics differ between the linear model from last week and the polynomial model we fit today?  Which model performs better on predicting the price of different packages of pumpkins?


**The root mean squared error (rmse) is significantly lower for our new model (poly_results) than in our old model (lm_results), at 3.27 and 7.23 respectively. The R-squared value for our new polynomial model is also much higher, at 0.89 indicating that 89% of the variation in our data can be explained by our model. Last week's R-squared value was only 0.49. The mean absolute error is also much lower in our polynomial model (2.34) than in our linear model (5.94) (units are price per package). This means that the polynomial model performs better on predicting the price of different packages of pumpkins.**

Let's visualize our model results.  First prep the results by binding the encoded package variable to them.
```{r}
# Bind encoded package column to the results
poly_results <- poly_results %>% 
  bind_cols(package_encode %>% 
              rename(package_integer = package)) %>% 
  relocate(package_integer, .after = package)


# Print new results data frame
poly_results %>% 
  slice_head(n = 5)
```

OK, now let's take a look! 

Question 4: Create a scatter plot that takes the poly_results and plots package vs. price.  Then draw a line showing our model's predicted values (.pred). Hint: you'll need separate geoms for the data points and the prediction line.
```{r}
# Make a scatter plot
poly_line_plot <- poly_results %>% 
  ggplot(mapping = aes(x = package_integer, y = price)) +
   geom_point(size = 1.6) +
   # Overlay a line of best fit
   geom_line(aes(y = .pred), color = "orange", linewidth = 1.2) +
   xlab("package")

poly_line_plot

```

You can see that a curved line fits your data much better.

Question 5: Now make a smoother line by using geom_smooth instead of geom_line and passing it a polynomial formula like this:
geom_smooth(method = lm, formula = y ~ poly(x, degree = 3), color = "midnightblue", size = 1.2, se = FALSE)

```{r}
# Make a smoother scatter plot 
poly_smooth_plot <- poly_results %>% 
  ggplot(mapping = aes(x = package_integer, y = price)) +
   geom_point(size = 1.6,
              alpha = 0.7,
              color = "midnightblue") +
   # Overlay a line of best fit
   geom_smooth(method = lm,
               formula = y ~ poly(x, degree = 3),
               color = "orange",
               size = 1.2,
               se = FALSE) +
   xlab("package")

poly_smooth_plot
  
```

OK, now it's your turn to go through the process one more time.
 
Additional assignment components :
6. Choose a new predictor variable (anything not involving package type) in this dataset.

I'm choosing variety.

7. Determine its correlation with the outcome variable (price).  (Remember we calculated a correlation matrix last week)

```{r}
# correlation between price and variety

# using baked pumpkins here because we want to find correlation (everything needs to be numeric)
cor(baked_pumpkins$variety,
    baked_pumpkins$price)

```


8. Create and test a model for your new predictor:
  - Create a recipe
  - Build a model specification (linear or polynomial)
  - Bundle the recipe and model specification into a workflow
  - Create a model by fitting the workflow
  - Evaluate model performance on the test data
  
```{r variety_model}
set.seed(123)

# we split our data into training and test data in Lab 1


# create a recipe ----
poly_pumpkins_recipe_variety <-
  recipe(price ~ variety, data = pumpkins_train) %>%
  step_integer(all_predictors(), zero_based = TRUE) %>% 
  step_poly(all_predictors(), degree = 3) # specifying polynomial


# build a model specification called poly_spec_variety ----
poly_spec_variety <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")

# bundle the recipe and model specification into a workflow ----
poly_wf_variety <- workflow() |> 
  add_recipe(poly_pumpkins_recipe_variety) |> 
  add_model(poly_spec_variety) 


# create a model by fitting the workflow ---- 
poly_wf_fit_variety <- fit(data = pumpkins_train, object = poly_wf_variety)

# ---- Evaluate model performance on test data ----

# make price predictions on test data 
poly_results_variety <- poly_wf_fit_variety %>% 
  predict(new_data = pumpkins_test) %>% 
  bind_cols(pumpkins_test %>% # bind multiple dfs by these columns
              select(c(variety, price))) %>% 
  relocate(.pred, # put prediction column after price column for easier comparison
           .after = last_col()) 

# evaluate how the model performed using yardstick::metrics()
variety_mod_performance <- metrics(data = poly_results_variety,
                                   truth = price, # evaluate based on true price
                                   estimate = .pred) # name of estimate column

variety_mod_performance


```
  
  
  - Create a visualization of model performance
  
Lab 2 due 1/24 at 11:59 PM
